Distributional Trees and Forests

### 2017-12-06 pkg/disttree/inst/CMStatistics2017/slides_rain.Rnw
small changes in formatting
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1827 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 31f5403ff6dc4a176afb5023f168f7c3208fa5f6

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber]{uibk}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Goal:} Predict total precipitation amount in alpine regions.\\
\medskip
\textbf{Input:} Numerical Weather Prediction (NWP):\\ 
Global physical model numerically simulating atmospheric processes on $50\times 50 \text{km}^2$ grid.\\
\medskip
\textbf{Problem:} Uncertain initial conditions, unresolved processes.\\ %$\rightarrow$ requires post-processing.\\
\vspace{1.3cm}
%\bigskip
\textbf{Distribution Assumption:}
%\vspace{0.1cm}
%\hspace{1cm}
$(\text{Total precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2)$\\
\bigskip
\textbf{Idea:} Statistical model with numerical predictions as covariates.
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Observation data} (National Hydrographical Service):\\
\begin{itemize}
\item 95 observation stations in Tyrol, Austria.\\
\item Daily 24h precipitation sums from July of 28 years (1985--2012).
\end{itemize}
\vspace{1cm}
%\bigskip
\textbf{NWP model output} (Global Ensemble Forecast System):\\
\begin{itemize}
\item Numerical ensemble predictions of various meteorological quantities.\\
\item 11 members each using slightly different perturbed initial conditions.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
\textbf{Base Variables:}
\begin{itemize}
\item Total precipitation
\item Convective available potential energy
\item Downwards short wave radiation flux (``sunshine'')
\item Mean sea level pressure
\item Preciptable water
\item 2m maximum temperature
\item Total column-integrated condensate
\item Temperature 
\item Temperature differences in altitude
\end{itemize}
\medskip
\textbf{Variations:} Ensemble min/max/mean/standard deviation\\
\vspace{0.1cm}
$\rightarrow$ 82 covariates.

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Application for one station:
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
Application for one station:
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item 10 times 7-fold cross validation.
\item Compare to commonly used heteroscedastic censored Gaussian models:
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize
\nodindent
\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
Application for all 95 stations:
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\item Compare to commonly used heteroscedastic censored Gaussian models:\\
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize
\nodindent
\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}
\end{frame}

@@

layout(cbind(1, 2), width = c(9, 1))
par(mar = c(5,4,4,0.1))
raster::image(tirol.dem, col = rev(gray.colors(100)),
              main="Observation Stations", ylab = "Latitude", xlab = "Longitude", 
              xlim = c(9.8,13.2), 
              ylim = c(46.6, 47.8), 
              cex.lab=1.5, cex.axis=1.2, cex.main=2)
plot(tirol.gadm, add = TRUE)
points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.9)
legend("topleft", pch=c(21,24,25,22), legend = c("Forest", "prespecified GAM", "boosted GAM", "EMOS"), cex = 1.2, bty = "n")

@@

Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
  \dquote{Calibrated probabilistic forecasting using ensemble model output statistics and minimum CRPS estimation}
  \textit{Monthly Weather Review}, 
  \textbf{133}(5), 1098--1118.
  \doi{10.1175/MWR2904.1}

@@

Hofner B, Mayr A, Fenske N, Schmid M (2017).
  \dquote{{gamboostLSS}: Boosting Methods for {GAMLSS} Models}
  \textit{{R} package version 2.0-0},
  \url{https://CRAN.R-project.org/package=gamboostLSS}

@@

Gebetsberger M, Messner J W, Mayr G J, Zeileis A (2017).
  \dquote{Fine-Tuning Non-Homogeneous Regression for Probabilistic Precipitation Forecasts: Unanimous Predictions, Heavy Tails, and Link Functions}
  \textit{Monthly Weather Review}, 
  \textbf{Forthcoming}.
  \doi{10.1175/MWR-D-16-0388.1}

>>>>

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber, url]{uibk}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Goal:} Predict daily precipitation amount in alpine regions.\\
\medskip
\textbf{Input:} Numerical Weather Prediction (NWP) based on a global physical model numerically simulating atmospheric processes on a $50\times 50 \text{km}^2$ grid.\\
\medskip
\textbf{Problem:} Uncertain initial conditions, unresolved processes.\\ %$\rightarrow$ requires post-processing.\\
\vspace{1.3cm}
%\bigskip

\textbf{Idea:} Statistical model with numerical predictions as covariates.\\
\bigskip
\textbf{Distribution assumption:}\\
\vspace{-0.3cm}
%\hspace{1.5cm}
%\center
$$(\text{precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2) \quad \text{censored at 0.}$$

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Observation data} (National Hydrographical Service):\\
\begin{itemize}
\item 95 observation stations in Tyrol, Austria.\\
\item Daily 24h precipitation sums from July over 28 years (1985--2012).
\end{itemize}
\vspace{1cm}
%\bigskip
\textbf{NWP model output} (Global Ensemble Forecast System):\\
\begin{itemize}
\item Ensemble predictions of various meteorological quantities related to precipitation, temperature,
air pressure, \dots.\\
\item 11 members each using slightly different perturbed initial conditions.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
\textbf{Base variables:}
\begin{itemize}
\item Total precipitation
\item Convective available potential energy
\item Downwards short wave radiation flux (``sunshine'')
\item Mean sea level pressure
\item Preciptable water
\item 2m maximum temperature
\item Total column-integrated condensate
\item Temperature 
\item Temperature differences in altitude
\end{itemize}
\medskip
\textbf{Variations:} Ensemble min/max/mean/standard deviation\\
\vspace{0.1cm}
$\rightarrow$ 82 covariates.

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
\textbf{Application for one station:}
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
\textbf{Application for one station:}
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item 10 times 7-fold cross validation.
\item Compare to commonly used heteroscedastic censored Gaussian models:
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize
\nodindent
\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
\textbf{Application for all 95 stations:}
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\item Compare to commonly used heteroscedastic censored Gaussian models. %:\\
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}

%\vspace{0.1cm}
%\begin{table}[t!]
%\footnotesize
%\nodindent
%\hskip-0.4cm\begin{tabular}{ l l l }
%\hline
%Model & Type & Variable Selection \\
%\hline
%\vspace{-0.15cm}
%\textbf{Forest} & recursive    & automatic \\ 
%                & partitioning &           \\ 
%\hline
%\vspace{-0.15cm}
%\textbf{prespecified GAM} & spline  & based on expert \\
%                          & in each & knowledge      \\
%\hline
%\vspace{-0.15cm}
%\textbf{boosted GAM} & spline  & automatic  \\
%                     & in each &            \\
%\hline
%\vspace{-0.15cm}
%\textbf{EMOS} & linear & total precipitation mean \\
%              &        & and standard deviation   \\
%\hline
%\end{tabular}
%\end{table}
\end{frame}

@@

layout(cbind(1, 2), width = c(9, 1))
par(mar = c(5,4,4,0.1))
raster::image(tirol.dem, col = rev(gray.colors(100)),
              main="Observation stations", ylab = "Latitude", xlab = "Longitude", 
              xlim = c(9.8,13.2), 
              ylim = c(46.6, 47.8), 
              cex.lab=1.5, cex.axis=1.2, cex.main=2)
plot(tirol.gadm, add = TRUE)
points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.9)
legend("topleft", pch=c(21,24,25,22), legend = c("Forest", "prespecified GAM", "boosted GAM", "EMOS"), cex = 1.2, bty = "n")

@@

Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
  \dquote{Calibrated Probabilistic Forecasting using Ensemble Model Output Statistics and Minimum CRPS Estimation}
  \textit{Monthly Weather Review}, 
  \textbf{133}(5), 1098--1118.
  \doi{10.1175/MWR2904.1}

@@

Hofner B, Mayr A, Schmid M (2016).
  \dquote{{gamboostLSS}: An R Package for Model Building and Variable Selection in the {GAMLSS} Framework}
  \textit{Journal of Statistical Software},
  \textbf{74}(1), 1--31.
  \doi{10.18637/jss.v074.i01}

@@

Gebetsberger M, Messner J W, Mayr G J, Zeileis A (2017).
  \dquote{Fine-Tuning Non-Homogeneous Regression for Probabilistic Precipitation Forecasts: Unanimous Predictions, Heavy Tails, and Link Functions}
  \textit{Monthly Weather Review}, 
  \text{Forthcoming}.
  \doi{10.1175/MWR-D-16-0388.1}
